---
title: "WUP_04"
author: "Jessica Villa"
date: "Summer 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
require (mosaic)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

```{r}
gssDataframe=read.csv("GSScsv.csv")
```
WORDSUM is a numerical variable  

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results
```{r}
boxplot(WORDSUM~DEGREE,data=gssDataframe)
```

From the box plot it appears that, the students in less than high school have a median less than that of the students in high school and in Bachelor.  

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Numerical Descriptive Results
```{r}
favstats(WORDSUM~DEGREE,data=gssDataframe)
```
From this table, it appears that the median of the graduate students is the greatest with a 11. It looks like there is a difference in the ability in the WORSUM category.  

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results
```{r}
model1 <- aov(WORDSUM~DEGREE,data=gssDataframe)
summary.aov(model1)
```
The p-value is 0.000236. This is because the p-value is less than 0.05, we reject the null hypothesis.   
```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Based on the graph, there is a difference between High School-Graduate and also between Less than High school and Graduate. This is because those levels do not cross or touch the zero.  

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

The degree that the person has makes a difference between their average WORDSUM score. The graduate has a higher median and mean. While, having an equal maximum. We rejected the null hypothesis after seeing the AOV results. There is a clear difference. Graduates get the highest average and LT high school gets the lowest average. In a way, it's based off of common sense. A graduate with more experience is expected to perform better. 

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
